\LARGE \bf Matrix Multiplication-Driven Repulsive Fields for Manipulator Kinematic Obstacle Avoidance

### 2024-03-19 Articles/IROS/root.tex
removed historic versions
Mentor Corrections: PAGE 1

%Our methodology employs occupancy voxel grids to create repulsive velocities. These velocities dynamically guide robots away from obstacles based on proximity, facilitating safe navigation. By adjusting repulsive forces in real time, our approach ensures robots maintain a safe distance from hazards, effectively integrating global and local planning strategies to navigate dynamic environments efficiently. This pragmatic solution leverages voxel-based spatial assessments to avoid potential collisions, ensuring pathfinding and obstacle avoidance.
%
%\alert{---}
%
%\subsection{Motion Planning}
%
%Motion planning is a process of finding a sequence of joint configurations that lead the manipulator from initial joints configuration to the goal configuration, while adhering to the constraints and optimization criteria. This is possible because the manipulator has a redundant number of joints~\cite{siciliano1990kinematic}. That means that it has more degrees of freedom (DOF) - joints, than it needs to execute the primary task, that is reaching the End Effector (EE) goal. Consequently, the robot can adopt different constrained joint configurations optimised according to the secondary task while performing the primary task. Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
%
%%Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence. This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task. Consequently, the robot can adopt different joint configurations optimised according to the secondary task while performing the primary task. Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
%%
%%The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}. It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with itself, static obstacles or other agents in the environment. 
%
%%In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
%
%There are global and local motion planning methods. Global, sampling-based, methods~\cite{vsvestka1997motion,lavalle1998rapidly, karaman2010incremental,kuffner2000rrt,gammell2015batch} offer a globally optimal solution based on a global search in configuration space. However, the generated trajectories are not smooth and optimal, and the performance of methods is insufficient for real-time operation.
%
%%There are global and local motion planning methods. Global, sampling-based, methods such as PRM~\cite{vsvestka1997motion}, RRT*~\cite{lavalle1998rapidly, karaman2010incremental} RRT-Connect~\cite{kuffner2000rrt}, Informerd RRT*~\cite{gammell2014informed}, BIT*~\cite{gammell2015batch} offer a globally optimal solution based on a global search in configuration space. However, the generated trajectories are not smooth and optimal, and the performance of methods is insufficient for real-time operation.
%
%Local motion planning methods employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c23}.  Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions. However, since they do not plan further than one step ahead, they tend to get stuck in local minima. 
%
%Often best solution to motion planning is a combination of a higher-level planner, for global static environment based path planning and local planner, that takes dynamic environment changes into account.
%
%%\subsection{Kinematic Obstacle Avoidance}
%%
%%When using inverse kinematics control for motion planning, one method is augmentation of task space to encompass supplementary parameters, in addition to the position and orientation of the end effector~\cite{c34, c35, c36}. This method returns a singular solution of a weighted combination of both tasks. Alternatively, our paper utilizes task prioritization~\cite{c29, c38, c41}, which involves calculating the null space for the primary task velocities and applying only those secondary velocities that do not directly influence the primary task velocities.
%
%\subsection{Artificial Potential Field}
%
%Khatib~\cite{c33} proposes the concept of an Artificial Potential Field. This field consists of a repulsive component that deflects the end effector away from obstacles, depicted as geometric primitives, and an attractive component that draws the manipulator towards its target. In the following years many different modifications and improvement of the original APF idea have been proposed, often focused on removal of local minimas in the potential field. Khim and Khosla~\cite{c40} suggested the use of harmonic functions to solve the problem of local minima. Pinto et al. \cite{c43} proposes to vary the field based on the distance of robot from obstacles to fill the local minima. Many researchers tried using APF as a heuristic to better guide sampling based approaches~\cite{c45, c46, c47}.
%
%Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator. Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}. Park et al.~\cite{park2020trajectory} used a numerical Jacobian in combination with APF for motion planning. 
%
%%Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima. 
%
%%Chen et al. proposes\remove{ an application }using APF in joint space and a variable kinematic optimization step~\cite{c50}. Long~\cite{c44} suggests creating motion plan of the manipulator using APF, he extends it using RRT to calculate virtual attractive point for the robot to move towards in case of local minima. Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
%%
%
%\subsection{ESDF creation}
%
%ESDF grids can be generated directly using sensor measurements. Oleynikova et al.~\cite{oleynikova2017voxblox} proposed a method for calculating the ESDF from TSDF. Han et al.~\cite{han2019fiesta} proposed a way to integrate point-cloud data into ESDF using ray-casting.  
%
%Another way is to generate ESDF from occupancy grids, that can be previously generated using sensors or based on predefined maps. When generating ESDF from occupancy grid, common approach is the Brushfire method~\cite{lau2010improved}, that spreads from obstacles until it calculates the distance for every field on a grid. Jump Flooding Algorithm (JFA)~\cite{rong2006jump} is a similar method, that can be implemented on a GPU for faster parallelized distance calculation.  
%
%Methods that allow for the dynamical distance field generation are rare. Zhou et al.~\cite{zhou2021egoplanner} propose use of pairs of points on trajectory and obstacles in their quadcopter trajectory optimization algorithm.
%
%%\remove{The field of image processing invented many different morphological algorithms, that work on binary images and calculate a value for every pixel in image, that represents the distance from nearest occupied binary pixel~\cite{szeliski2021computer, felzenszwalb2012distance, rosenfeld1968distance}. Approaches can be seperated based on the distance metric that they use. }
%
%%Euclidean Signed Distance Fields (ESDFs) are valuable in robotic navigation and 3D reconstruction, providing a measure of the shortest distance to the nearest obstacle for each point in space. The creation of ESDF grids can leverage sensor measurements directly or be derived from other data structures like Truncated Signed Distance Fields (TSDFs) or occupancy grids.
%%
%%One prominent method for generating ESDFs is converting TSDFs into ESDFs. Oleynikova et al. [(2017)] introduced a technique that efficiently computes the ESDF from TSDFs, utilizing incremental updates for dynamic environments. This approach is significant for real-time applications in robotics, where rapid updates to the environment model are crucial. Similarly, Han et al. [(2019)] developed a method that integrates point-cloud data into ESDF using ray-casting, offering a way to construct ESDFs directly from sensor data.
%%
%%Alternatively, ESDFs can be generated from occupancy grids, which themselves are often derived from sensor measurements. The Brushfire algorithm, as improved by Lau et al. [(2010)], is a commonly used technique in this context. It operates by spreading from obstacles within the grid, calculating the minimum distance to an obstacle for each grid cell. This approach is particularly effective for static environments where the layout of obstacles does not change frequently.
%%
%%Furthermore, the Jump Flooding Algorithm (JFA) offers another method for generating ESDFs from occupancy grids. JFA is notable for its suitability for implementation on GPUs, enabling faster parallel distance calculations. This advantage is critical in applications requiring rapid processing of large datasets, such as in high-resolution 3D mapping and complex robotic navigation tasks.
%%
%%In conclusion, the generation of ESDFs from sensor measurements or derived data structures like TSDFs and occupancy grids is a crucial aspect of robotic navigation and 3D environment modeling. Techniques like those proposed by Oleynikova et al. [(2017)], Han et al. [(2019)], and the improvements to the Brushfire algorithm by Lau et al. [(2010)], as well as the application of JFA, highlight the ongoing advancements in efficient and accurate ESDF creation methodologies.
%
%
%
%\alert{---}
%
%In dynamic environments, a robot manipulator must swiftly respond to emerging obstacles. Our approach enables the robot to determine safe movement directions and velocities by real-time assessment of nearby space for obstacles.
%
%In 3D graphics, a voxel is like a tiny cube in a grid that makes up the 3D space. We use voxels to represent the workspace of the robot, each voxel has a value that tells us if it’s probabilistically empty or occupied. A zero means empty, and a one means definitely something is there, with values in between showing varying levels of certainty.
%
%%Mreža voxels je lahko predefinirana, glede na model / 3d zemljevid prostora. Zasedenost voxlov lahko spreminjamo glede na poznane pozicije in trajektorije preostalih agentov v prostoru. Kot omenjeno v uvodu, pa lahko zasedenost voxlov pridobimo tudi z senzorskimi sistemi. 
%
%The occupancy grid of voxels can be set in advance to match the layout of the space. We can also change the values in the voxels if we know where other objects or agents are moving. Besides pre-setting, we can use depth sensors to detect what's around.
%
%Our method calculates repulsive velocities from existing occupancy voxel grid, which are vectors that guide the robot on where to go by pointing away from potential collisions. These vectors gain strength as the robot approaches an obstacle, signaling an urgent need to maneuver to safety. Conversely, when the robot is positioned at an equal distance from obstacles in the vicinity—safely away from all potential collisions—the repulsive velocities weaken because there’s no immediate need to move. 
%
%%In our method, we compute repulsive velocities within the task space using a novel matrix kernel multiplication approach. Concentrating on the task space is advantageous as it provides a more direct and realistic representation of the environment. 
%
%%\replace{Naša metoda je posebno primerna za uporabo z senzorskimi sistemi kot so LIDAR ali globinske kamere, saj zaradi upoštevanja celotne okolice točke in ne le razdalje do najbližje točke v okolici efektivno filtriramo senzorski šum.}{prestavi to v diskusijo}
%
%Repulsive velocities tell the agent in which direction to move, so that it avoids nearby obstacles. These velocities drop to zero when the agent maintains a minimum safe distance from obstacles, and rise to their highest when it nears an obstacle, facilitating immediate evasive action. As the repulsive field calculation is locally based, it will also go to zero when the agent is surrounded by all directions, equally spaced from all sides. That is, it is in the best local minima away from all the obstacles in its proximity.

@@

This section introduces our novel approach for calculating repulsive fields, a critical component in enabling robots to detect and avoid obstacles in real-time. By leveraging a voxel-based representation of the surrounding space, our method dynamically assesses potential collision threats and calculates directional repulsive velocities. These velocities are essential for guiding the robot away from obstacles, ensuring smooth and safe navigation through complex environments. Our innovative use of kernel convolution methods allows for the precise calculation of avoidance velocities, providing a robust solution for obstacle avoidance that can be integrated seamlessly into robotic control systems. The following subsections delve into the technical details of our approach, including the mapping of Cartesian space to a discrete occupancy grid, kernel selection for optimized field calculation, and the application of tri-linear interpolation for smooth velocity transitions.

%\subsection{MAPPING}

%The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous, as rounding the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance.\add{is this clear, do i need equation?} However, this discretization can sometimes lead to discontinuities. Therefore, we propose a second approach: tri-linear interpolation of the calculated repulsive field to achieve a continuous repulsive field value.

\subsection{KERNELS AND THEIR MATRIX APPLICATIONS}
%\subsection{Selecting Kernels and Matrix Operations}

The fundamental concept of our directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.

%\remove{Our filters structure was inspired by the Sobel  operator, a 2D convolutional filter frequently utilized in computer vision for calculating image gradients at specific points.}

Our kernels are designed as three-dimensional matrixes with a primary kernel axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity. The two secondary kernel axes are orthogonal to this primary axis. The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where jump between max positive and max negative magnitude is. The function of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity changes as the agent approaches an obstacle. Moreover, it is essential for the magnitudes at the kernel's periphery to be minimal, promoting a smooth increase in repulsive velocity when approaching the obstacle rather than a sudden spike.

% KERNEL WEIGHTS

% PRIMARY AXIS

%We propose two different primary axis weights distributions, of course there is no reason why any other distribution of weights could not be used. The choice of the weights distribution should depend on the profile of the repulsive velocities we want to archieve for the APF. 
%
%The first of the proposed functions is a mirrored normal / gaussian distribution. By changing the sigma we can control how fast or slow does the field value grow when we approach obstacles.
%
%$\Delta i = c_i - i$, $\Delta j = c_j - j$ in $\Delta k = c_k - k$ predstavljajo število celic odmika od centralnega polja matrike, v katerem se nahaja naša točka na agentu v posamezno koordinatno smer.

We introduce two distinct distributions for the primary axis weights, although other distributions could also be applicable. The selection of weight distribution is contingent on the desired profile of repulsive velocities within the Artificial Potential Field (APF).

The inaugural proposed function adheres to a mirrored normal or Gaussian distribution, wherein adjusting the standard deviation (\(\sigma\)) modulates the rate at which field values escalate as obstacles are approached.

The terms \(\Delta i = c_i - i\), \(\Delta j = c_j - j\), and \(\Delta k = c_k - k\) denote the displacement count from the matrix's central field, where the field components are calculated, in their respective coordinate directions.

@@

%\begin{equation}
%w_i = 
%\begin{cases} 
%	\frac{e^{-\frac{\Delta i^2}{2\sigma^2}}}{\sigma \sqrt{2\pi}} & \text{if } \Delta i > 0 \\
%	0 & \text{if } \Delta i = 0 \\
%	-\frac{e^{-\frac{\Delta i^2}{2\sigma^2}}}{\sigma \sqrt{2\pi}} & \text{if } \Delta i < 0 
%\end{cases}
%\end{equation}

@@

% ORTHAGONAL AXIS

@@

%\remove{In the orthogonal axes, the magnitude of values diminishes with increasing distance from the core of the kernel, nearing zero at the edges. This gradation is vital for a smooth velocity transition when obstacles emerge in the kernels range. The peak magnitude is located at the kernel's core, along the primary axis. The polarity of the values is determined by their position relative to the primary axis. }

@@

%Once the weights of the repulsive kernels have been generated, we can use kernel multiplication to get the repulsive velocity components. This method is tailored to calculate the components of the avoidance velocity vector in the Cartesian space, we generate the corresponding kernel and extract a segment of the obstacle grid of matching size, centered at agent or the point of interest (POI), resulting in two same size 3D matrices—one is a "window" from the obstacle grid $A_d$ and the other representing the kernel $K_d$.

@@

%If the agent is located near the edge of our known voxel grid we can set the elements of the window would be located in the space beyond the matrix as empty in which case the robot might want to move towards this space or as occupied, which will prevent the robot from moving out of the known grid space.

@@

%By employing the Hadamard (element-wise) product (eq.~\ref{eq:matrix-product}) between the cutout segment of the obstacle grid \(G_d\) and the corresponding 3D convolutional kernel \(W_d\) and than summoning all of the matrix values for each of the directions $d \in \{x, y, z\}$, we derive the resultant repulsive velocities vector \( \vec{v}_i = \left[ v_{x_i} v_{y_i} v_{z_i}\right] \).

@@

%\begin{equation}
%	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
%	\label{eq:summation}
%\end{equation}
%

@@

%The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous. In the transition from Cartesian space coordinates to discrete occupancy grid representations, approximating or rounding can induce discontinuities in the resulting repulsive field. To mitigate this effect, we employ tri-linear interpolation, enabling a seamless transition to a continuous repulsive field value (Fig. \ref{fig:interp-experiment}).
%
%It is essential that the velocity contributions affecting the robot change smoothly. However, since our obstacle grid is discretely defined, achieving perfect continuity can be challenging. Increasing the resolution of the obstacle field can theoretically bring us closer to continuous behavior, but in practice, we are constrained by finite resolution of our sensors and of the grid. Poleg tega bo višja resolucija posledično povzročila povečanje spomina potrebnega za shranjevanje zemljevida / mreže okolja in potrebne procesorske moči. To ensure that the velocity remains continuous when transitioning from one cell of the obstacle grid to another at a point of interest (POI), we employ trilinear interpolation.

@@

%\begin{equation}
% X = \left[ \lfloor p_{x} \rfloor, \lceil p_{x} \rceil \right] 
%\end{equation}
%
%\begin{equation}
% Y = \left[ \lfloor p_{y} \rfloor, \lceil p_{y} \rceil \right] 
%\end{equation}
%
% \begin{equation}
% Z = \left[ \lfloor p_{z} \rfloor, \lceil p_{z} \rceil \right] 
%\end{equation}

@@

%\begin{equation}
%	\Delta X = \frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( X(2) - X(1) \right)}
%\end{equation}
%
%\begin{equation}
%	\Delta Y = \frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Y(2) - Y(1) \right)}
%\end{equation}
%
%\begin{equation}
%	\Delta Z = \frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Z(2) - Z(1) \right)}
%\end{equation}

@@

%\begin{equation}
%	\vec{V}_{rep,00} = \vec{V}_{rep,000}(1 - x_d) + \vec{V}_{rep,100}x_d
%\end{equation}
%\begin{equation}
%	\vec{V}_{rep,01} = \vec{V}_{rep,001}(1 - x_d) + \vec{V}_{rep,101}x_d
%\end{equation}
%\begin{equation}
%	\vec{V}_{rep,10} = \vec{V}_{rep,010}(1 - x_d) + \vec{V}_{rep,110}x_d
%\end{equation}
%\begin{equation}
%	\vec{V}_{rep,11} = \vec{V}_{rep,011}(1 - x_d) + \vec{V}_{rep,111}x_d
%\end{equation}

@@

%\begin{equation}
%	\vec{V}rep_{xyz,0} = \vec{V}rep_{xyz,00}(1 - \Delta y) + \vec{V}rep_{xyz,10} \, \Delta y
%\end{equation}
%
%\begin{equation}
%	\vec{V}rep_{xyz,1} = \vec{V}rep_{xyz,01}(1 - \Delta y) + \vec{V}rep_{xyz,11} \, \Delta y
%\end{equation}

@@

%The final result is a smooth repulsive velocity field that transitions smoothly between the discrete values calculated at distinct points in the obstacle grid.

@@

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} % Replace 'example-image' with the filename of your image
	\caption{Repulsive field generation via convolutional kernels. Left: Original repulsive fields in X (top) and Y (bottom) directions. Right: Corresponding fields using interpolation, showcasing enhanced smoothness.}
	\label{fig:interp-experiment}
\end{figure}

@@

%\begin{equation}
%	\dot{\vec{q}} = J^+ \xi_{p} \mathrm{\vec{v}_{att}} + \dot{\vec{q}}_{rep}
%\end{equation}

@@

%\remove{For redundant robots, it optimizes solution selection by minimizing the joint velocity norm, while damping is applied to limit excessive joint velocities near singular configurations, thus improving solution robustness.}

%\add{maybe add the middle joints component in the equation (it is not necessary, doesn't make much change in the chosen experiments)}

@@

%Vodenje vrha manipulatorja (end-effector) je naloga z najvišjo prioriteto.

@@

%In the above equation (eq.~\ref{eq: v_att}), $\vec{v}$  represents the end effector's translational velocity towards the target, combining direction and magnitude. The terms $\vec{x}_{EE}$ and $\vec{x}_g$ denote the current and goal positions of the EE, respectively, in Cartesian coordinates. The unit vector calculation, $\frac{\vec{x}_{EE}-\vec{x}_g}{||\vec{x}_{EE}-\vec{x}_g||}$, ensures motion directed towards the target. Finally, the sigmoid function, particularly the arctangent component, modulates this velocity to avoid overshooting, balancing speed and precision. The EE velocity will be transformed using the sigmoid curve when near the goal. When the distance from goal point is high, the sigmoid will cause the velocity to get into saturation, that is upper limit. The constant $k_{sigm}$ allows us to set how close to the goal does the robot EE start slowing down and what is the velocity profile when approaching the goal point.

@@

%The rotational error of the EE is needed for ensuring goal orientation of the EE. In our approach, orientations are depicted using rotation matrices. Specifically, \( R \) represents the current EE orientation, while \( gR \) signifies the goal EE orientation. The disparity between these orientations is encapsulated by the relative rotation matrix \( dR \). This matrix is formulated by multiplying the goal orientation matrix \( gR \) with the transpose of the current orientation matrix \( cR^{T} \). To ensure that it represents a pure rotation without any scaling we than normalize the so gotten matrix .

%Determination of the requisite rotation adjustment utilizes the relative rotation matrix \(dR\), calculated from the multiplication of \(gR\) by the \(cR\) transpose, and subsequent normalization to ensure a pure rotational representation. This method guarantees that \(dR\) accurately captures the orientation discrepancy, devoid of any scaling factor.

%The current orientation of the end-effector (EE) and its goal orientation are encoded through rotation matrices, \(cR\) for the current state and \(gR\) for the goal state. The necessary rotational adjustment is identified through the relative rotation matrix \(dR\), generated by multiplying \(gR\) with the transpose of \(cR\) and then normalizing the outcome (eq. \ref{eq: rot_diff_mat}). This process ensures \(dR\) accurately represents rotational differences, without any scaling factors.
%
%To enhance computational efficiency and prevent gimbal lock, the approach transitions from matrix to quaternion representations for precise and efficient rotational velocity adjustments, essential for achieving the target orientation of the EE.

@@

%\begin{equation}
%	dR = \frac{gR \cdot cR^{T}}{||gR \cdot cR^{T}||}
%	\label{eq: rot_diff_mat}
%\end{equation}

%\begin{equation}
%	\boldsymbol{dR} = \frac{\boldsymbol{gR} \cdot \boldsymbol{cR}^{T}}{\left\lVert\boldsymbol{gR} \cdot \boldsymbol{cR}^{T}\right\rVert}
%	\label{eq: rot_diff_mat}
%\end{equation}

@@

The relative rotation matrix is converted into a quaternion, which is subsequently logarithmically transformed. The resulting quaternion's components, excluding the real part, constitute the rotational error vector \( \vec{\omega} \).
%
%\begin{equation}
%	dR \mapsto dQ = a + b \, i + c \, j + d \, k
%	\label{eq: quat_mapsto}
%\end{equation}
%
%\begin{equation}
%	dQl = 2 \cdot \log(dQ) = al + bl \, i + cl \, j + dl \, k
%	\label{eq:quat_log}
%\end{equation}

@@

%Ker uporabljamo preslikavo v ničelni prostor primarne hitrosti (null space), lahko v primeru velikih hitrosti primarne naloge preostane premalo prostostnih stopenj, da bi se lahko varno izognili oviram. Torej sekundarna naloga izogibanja nima dovolj sposobnosti, prostosti za delovanje.
%
%To ensure the primary task doesn't overpower the secondary task, we've integrated a primary task execution slowdown. This mechanism can reduce the manipulator's velocity towards its primary goal, leaving more maneuverability space for the secondary tasks.
%
%\begin{equation}
%	\label{eq:slowdown}
%	\xi_{p}=
%	\frac{1}{1 + \kappa_{\text{sec}} v_{\text{Rmax}}}
%\end{equation}
%
%The slowdown factor (eq.~\ref{eq:slowdown}) is influenced by the constant \( \kappa_{\text{sec}} \) and the maximum repulsive velocity vectors magnitude, which depends on robot's minimum distance from an obstacle \( v_{\text{Rmax}} \). We calculate the minimal distance in the repulsive velocities phase of the algorithm, as explained in section~\ref{chap:repulsive velocity}. As per the equation, a small \( v_{\text{Rmax}} \) minimizes the slowdown effect, allowing uninterrupted primary task execution. Conversely, a large \( v_{\text{Rmax}} \) increases the slowdown, by making the $\xi_{p}$ factor smaller, giving the secondary task more time for corrective actions.

@@

%To move the manipulator away from the obstacles in its environment, we cover the manipulator evenly with virtual detection points (POI). Gostota virtualnih točk je odvisna od velikosti matrik. Če je $w$ širina matrike v smeri pravokotno na primarno smer, je dobro, če so točke medsebojno razmaknjene glede na potek segmentov za $\Delta = \frac{w}{3}$. Tedaj se bodo posamezne matrike deloma pokrivale in dosežemo enakomerno porazdelitev točk (POI) po celotnem robotu. Med izvajanjem kinematične optimizacije izračunavamo izogibne / odbojne hitrosti za vsako od točk po postopku opisanem v zgornjem poglavju. \add{referenca poglavja} To dobimo tako, da izvedemo preslikavo vsake od točk v task space, kjer izvajamo izračun odbojne hitrosti. Pri tem je posamezna kartezična preslikava sestavljena iz kinematične preslikave iz baze robota do začetka segmenta na katerem se točka nahaja in dodane preslikave do izbrane točke (od začetka segmenta do dela kjer se nahaja točka). \(T_{0 \rightarrow \text{POI}} = T_{0 \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}\) Za vsako od točk dobimo tri komponente kartezične hitrosti.

@@

%\begin{equation}
%	T_{0 \rightarrow \text{POI}} = T_{0 \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}	
%	\label{eq:transformations}
%\end{equation}

@@

%Ko dobimo za vsako od točk odbojno hitrost (repulsive / avoidance velocity), izračunamo drugo normo vsake od hitrosti. Za $K$ točk, v katerih so izogibne hitrosti največje in so posledično najbližje oviri nato izračunamo sklepne hitrosti, ki povzročijo izogibanje oviri.

@@

%\begin{equation}
%J_{d_0}^+ = N J_{d_0}' (J_{d_0} N J_{d_0}' + \sigma_{rep})^{-1};
%\end{equation}

%\begin{equation}
%	J_{d_i}^{+} = N J_{d_i}' (J_{d_i} N J_{d_i}' + \sigma_{rep} I)^{-1}, \quad \text{for } i = 1, 2, \ldots, K;
%\end{equation}

@@

%Za vsako od izbranih točk na robotu izračunamo Jacobijevo transformacijo hitrosti, ki preslika translacijske hitrosti izbrane točke na robotu iz kartezičnega prostora, v prostor sklepov na robotu, ki povezujejo bazo z izbrano točko. V izbranih točkah nas ne zanimajo transformacije napak orientacij. V resnici nas zanima samo komponenta translacijske hitrosti v smeri normale, ki kaže od najbližje točke na objektu proti izbrani točki interesa (POI) na manipulatorju.

%We calculate translational velocity Jacobian for every of the $K$ selected points on the robot. Since we are only interested in the part of velocity that leads us directly away from the obstacles, we can

%
%Naš prostor operacij se posledično skrči na eno dimenzijo in the Jacobian, which relates the joint space velocities q and the velocity in the direction of do, can be calculated as  $J_{d_i} = \vec{n}_i^T J_i$.  \add{citat:žlajpah} Pri čemer je vektor normale želene avoidance velocity kar enotski vektor hitrosti, ki ga dobimo s pomočjo naših matrik $\vec{n}_{i} = \frac{\vec{v}_{poi i}}{||\vec{v}_{poi i}||}$. This method offers computational efficiency by eliminating complex matrix inversions, as the Jacobian changes from matrix into a vector dimensions $1 \times n$, where n is number of joints that are located from before the POI and simplifying repulsive velocity calculations.

@@

%Celotne izogibne sklepne hitrosti dobimo z obteženo vsoto izogibnih sklepnih hitrosti izbranih $K$ najbližjih točk, pri čemer upoštevamo vpliv primarne naloge gibanja vrha manipulatorja na sklepne hitrosti.

@@

%Prvo predstavimo rezultirajoče se odbojno polje za 2D - dvo dimenzijski zemljevid. Načeloma je delovanje na treh dimenzijah enako, seveda pa je to težje vizualizirati. Pri tem smo uporabili našo metodo izračuna odbojnega potencialnega polja hitrosti. Namesto uporabe v točkah robotskem manipulatorju smo točke vzorčili preko celotnega zemljevida ovir, kot je vidno na figur \ref{fig:example-field}. V vsaki od vzorčenih točk smo izvedli v članku opisani postopek pridobivanja odbojnega polja v X in Y smeri. Na spodnjem prikazu figure smo nato vizualizirali odbojno poljo kot vektorsko polje sestavljene iz X in Y komponente.

@@

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
	\caption{Visualization of the potential field calculated for the entire map using the proposed approach. Top: Obstacle distribution via a 2D heatmap. Center: Induced repulsive fields along X (left) and Y (right) axes. Bottom: Composite vector field showcasing the resultant repulsive velocities for obstacle avoidance.}
	\label{fig:example-field}
\end{figure}
%
%\add{add label, caption, every plot has different axis scale}
%
%\comment{}{we dont have the close obstacles bottlenecks, we do have weird bottlenecks in the orthagonal directions - maybe}
%
\subsection{Manipulator Examples}

%Delovanje potencialnega polja smo prikazali na dveh različnih primerih manipulatorja. Pri tem je ločljivost naše mreže voxlov ovir $R=10cm$. Uporaba interpolacije nam omogoča, da imamo relativno grobo mrežo voxlov, kar zniža spominsko zahtevnost mreže prostora, ter pospeši izračun. V primerih smo izbrali $K = 7$ POI točk interesa, v katerih gledamo oddaljenost od ovir, ki so enakomerno razporejene po segmentih in sklepih od drugega sklepa robota do vrha manipulatorja, torej od prve točke naprej, kjer se je manipulator sposoben izogniti oviri. POI so na grafih prikazani s pikami na manipulatorju. V bližini ovir iz njih segajo vektorji, ki prikazujejo izračunane odbojno hitrost v točkah.

%Med izvajanjem simulacije izvajamo eulerjevo integracijo izračunanih sklepnih hitrosti, s korakom $T_{step}=0.1s$.

@@

%V prvem primeru \ref{fig:keyframes-3d-column} manipulator z uporabo potencialnega polja izračunanega na predlagani način, se manipulator varno "zvije" oziroma izbere pot do točke, ki se nahaja na drugi strani stebra. Izbrane konstante primarne naloge so $k_v=5$ in $k_w=20$. Izbrane konstante obteženosti posameznih POI so $\alpha=\frac{\left[ \frac{3}{9}, \frac{2}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}
% \right]}{10}$. Najkrajša pot vrha v točko pelje direktno skozi steber. Primarna naloga je, da vrh preide v želeno pozicijo in orientacijo. Ker je naloga vrha primarna, moramo uvesti znižanje primarne hitrosti v bližini ovire $\xi_{p}=1$, da ima sekundarna naloga dovolj prostosti in časa, da lahko izvede rekonfiguracijo za obhod stebra. Simulacijo izvajamo $50$ korakov.

@@

%\bibitem{siciliano2010robot} B. Siciliano, L. Sciavicco, L. Villani, and G. Oriolo, Robot. Model. Plan. Control, 2010, pp. 161–189. [Online]. Available: http://link.springer.com/10.1007/978-1-84628-642-1\_4

>>>>

@@

We leverage a voxel-based representation of the surrounding space, to dynamically assesses potential collision threats and calculates directional repulsive velocities for smooth and safe navigation through complex environments.

\subsection{KERNELS AND THEIR MATRIX APPLICATIONS}

The fundamental concept of directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.

Kernels are designed as three-dimensional matrixes with a primary kernel axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity. The two secondary kernel axes are orthogonal to the primary axis. The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where the values jump from max negative to max positive magnitude (Fig.~\cite{fig:kernels}). The function of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity changes as the agent approaches an obstacle. Moreover, it is essential for the magnitudes at the kernel's periphery to be minimal, promoting a smooth increase in repulsive velocity when approaching the obstacle rather than a sudden spike.

We introduce two distributions for the primary axis weights. The selection of weight distribution is contingent on the desired profile of repulsive velocities within the Artificial Potential Field (APF).

First proposed function is mirrored normal or Gaussian distribution, the standard deviation (\(\sigma\)) modulates the rate at which field values escalate as obstacles are approached.

Terms \(\Delta i = c_i - i\), \(\Delta j = c_j - j\), and \(\Delta k = c_k - k\) denote the displacement index from the matrix's central field, where fields' components are calculated.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} 
	\caption{Repulsive field generation via convolutional kernels. Left: Original repulsive fields in X (top) and Y (bottom) directions. Right: Corresponding fields using interpolation, showcasing enhanced smoothness.}
	\label{fig:interp-experiment}
\end{figure}

@@

@@

@@

@@

@@

@@

@@

The relative rotation matrix is converted into a quaternion, which is subsequently logarithmically transformed. The resulting quaternion's components, excluding the real part, constitute the rotational error vector \( \vec{\omega} \).

@@

@@

@@

@@

@@

@@

@@

@@

@@

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
	\caption{Visualization of the potential field calculated for the entire map using the proposed approach. Top: Obstacle distribution via a 2D heatmap. Center: Induced repulsive fields along X (left) and Y (right) axes. Bottom: Composite vector field showcasing the resultant repulsive velocities for obstacle avoidance.}
	\label{fig:example-field}
\end{figure}

\subsection{Manipulator Examples}

@@

@@
